Nearest Neighbour Methods for Length of Stay Prediction

### 2015-02-09 papers/paper.tex
Cut out some of the results and discussion section

\subsection{Comparison of Feature Selection Methods}
The best accuracy and AUC results for
each classifier resulted from a reduced feature set obtained through a
particular feature selection method. There was at least one feature selector
for each classifier that managed to outperform the same classifier with
no feature selection. This is likely due to the large number of features in
the original data -- after cleaning and preliminary removal of redundant
features, we were still left with over 50. Many of these would still likely
be irrelevant to predicting the class.
Out of 20 combinations of classifiers
and discretisation (the 10 classifiers with and 10 without
discretisation), the C4.5 wrapper method resulted in the best
accuracy for 7 of these combinations, followed by 5 using information gain
and 4 using the correlation coefficient and very few for the other methods.
Under no combination was 1R the best feature selection method to use, as other
methods produced better accuracy.

@@

Note that regardless of whether we consider accuracy or AUC to be more
important in evaluating how our classifiers performed, neither the baseline
feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
the domain expert (both of which were manually selected) yielded better
results than automatic feature selection in all but two cases. We did find
that expert selection produced the best accuracy result for 1R (equal with
correlation coefficient selection, but used less features) and that the
baseline feature set produced the best accuracy for Na\"{i}ve Bayes. Neither
feature selection method produced the best AUC result for any classifier.
This is a surprising result because intuitively and in the literature it
has been stated that in a highly specialised field such as medicine, the domain
experts would be best-equipped to make judgements about which features to
use \cite{Witten2005}, and these should give reasonable results. However,
our results show that in attempting to classify patients according to their
LOS, we should consider both manual and automatic approaches.

Recall that $k$-NN with 1 nearest neighbour and C4.5 wrapper feature
selection produced the best accuracy (77.81\%) for the trauma data set, and
logistic regression with correlation coefficient feature selection at a
threshold of 0.1 gave the best AUC (0.846). These two feature selection
methods selected 11 and 29 features respectively, and the features that
were chosen in common between them were: \texttt{operation}, \texttt{bp},
\texttt{iss}, \texttt{lowerlimbnopelvis}, \texttt{age}, \texttt{lowerlimbany}
and \texttt{icu}. This adds evidence to the work of Dinh et al. who found
that \texttt{iss}, \texttt{age} and \texttt{operation} were important
distinguishers of LOS $\leq$ or $>$ 2 days \cite{Dinh2013a}. We also
confirm the findings of Gabbe et al. who found that \texttt{age},
\texttt{iss} and \texttt{bp} were useful in predicting the length of
hospital stay of blunt trauma patients \cite{Gabbe2005}. Our feature
selection results also
support the findings of McGonigal et al. \cite{McGonigal1993} who used
\texttt{age} and
\texttt{iss} to predict the probability of survival for trauma patients, a
related problem in the same medical domain.
Our observations suggest the use of feature selection not only as a
pre-processing step to improve the accuracy or AUC of a classifier, but also
as a tool used separately that could help extract insights from medical data.

Although we were able to achieve higher accuracy and AUC than the baseline,
our best approaches also have a number of limitations which need to be pointed
out. The first point to note is that using a wrapper method of feature
selection is slower than using a filter approach (or manually picking the
features before training), as the learning algorithm used as the wrapper
needs to be invoked at each of the folds of cross-validation. Additionally,
the $k$-NN algorithm becomes slower as the data set increases in the number
of examples. We did not face such a problem during our work, but this will
have to be addressed for larger data sets. It is also common to use more
than one nearest neighbour, so care needs to be taken to find a suitable number
of neighbours for a given problem. Finally, we achieved an improvement
in AUC of 0.034 but this was with 29 features, which is still a fairly large
number. The importance of using a fewer number of features is twofold: not
only does this speed up the training time, it also means that we are able to
make predictions without requiring a lot of information about a patient,
which is a crucial consideration if this is to be implemented for use by
physicians.

\subsection{Comparison of Classifiers}
Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
accuracy and AUC of all classifiers under each feature selection method for
the trauma data set.
The methods that do not select features based upon a specified threshold
(manual methods as well as CFS, C4.5 wrapper and no feature selection)
are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
based upon thresholds (correlation, information gain and 1R) are graphed
in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
threshold is changed.

In terms of classification accuracy, we did not find a lot of variation
between all classifiers, with Zero-R producing 66.91\% accuracy and 77.81\%
being the best we achieved -- an improvement of just over 10\%. The AUC of
the 11 classifiers we tested ranged from 0.498 with Zero-R to 0.846 with
logistic regression. Logistic regression, the \textit{de facto} method
used in LOS classification, performed consistently better than the others
across all feature
selection methods, but only when the feature set was not heavily reduced by
the feature selector: in these cases, such as the results from Figure
\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
selected by CFS, C4.5 decision trees and nearest
neighbour methods have higher accuracy than logistic regression. This is also
apparent from the graphs in Figure \ref{fig:tr-threshold}: for all three
feature selection methods, as we increase the threshold, logistic regression
drops in performance after the initial few thresholds and is outperformed by
C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
neighbours and K*), as well as our Ranked Distance algorithm.
This is true for both accuracy and AUC.

We found that in most cases, the SVM did not achieve a higher accuracy or
AUC than logistic regression with the same features even though SVMs are
considered the state-of-the-art in classification algorithms. Additionally,
as the feature set was reduced, the discriminating ability of SVMs (as
indicated by the AUC) decreased more than their accuracy: this is particularly
noticeable in Figure \ref{fig:tr-threshold-oner}, where the AUC of the SVM
drops sharply to below the AUC of even the 1R classifier at the 67.905\%
threshold, where there are
only 4 features. SVMs have not been used in a LOS classification problem like
ours, so there will need to be further work carried out in order to assess the
suitability of using SVMs to classify patient LOS.

@@

Nearest neighbour approaches appear to have performed well on this data set in
both accuracy and AUC, especially in comparison to more sophisticated
classifiers and when the feature set is significantly reduced. From the graphs
in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours
approaches perform similarly when the number of features is decreased, but our
Ranked Distance algorithm showed a statistically significant improvement over
other NN algorithms. This
improvement is not apparent at smaller feature sets, where using the standard
$k$-NN algorithm with 1 or 20 neighbours performs slightly better.
The performance of NN we observed in our results is
an interesting finding because nearest neighbours have not been commonly
applied in predicting the LOS, but there is intuitive appeal in the
case-by-case nature of how these algorithms classify new examples that should
be explored further. Unlike MLPs and SVMs, prediction decision of nearest
neighbours are more transparent and understandable. However, the major drawback
of NN algorithms is their storage requirement and the computational time
required to search for the nearest neighbours to make the classification.

>>>>

\subsection{Comparison of Feature Selection Methods}
The best accuracy and AUC results for
each classifier resulted from a reduced feature set obtained through a
particular feature selection method. There was at least one feature selector
for each classifier that managed to outperform the same classifier with
no feature selection. This is likely due to the large number of features in
the original data: after cleaning and preliminary removal of redundant
features, we were still left with over 50. Many of these would still likely
be irrelevant to predicting the class.
Out of 20 combinations of classifiers
and discretisation (the 10 classifiers with and 10 without
discretisation), the C4.5 wrapper method resulted in the best
accuracy for 7 of these combinations, followed by 5 using information gain
and 4 using the correlation coefficient and very few for the other methods.
\texttt{TODO: recount the numbers in this paragraph.}

@@

Note that regardless of whether we consider accuracy or AUC to be more
important in evaluating classifiers, neither the baseline
feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
the domain expert (both of which were manually selected) yielded better
results than automatic feature selection in all but two cases. Neither
feature selection method produced the best AUC result for any classifier.
This is a surprising result because intuitively and in the literature it
has been stated that in a highly specialised field such as medicine, the domain
experts would be best-equipped to make judgements about which features to
use \cite{Witten2005}, and these should give reasonable results. However,
our results show that in attempting to classify patients according to their
LOS, we should consider both manual and automatic approaches.

Recall that $k$-NN with 1 nearest neighbour and C4.5 wrapper feature
selection produced the best accuracy (77.81\%) for the data set, and
logistic regression with correlation coefficient feature selection at a
threshold of 0.1 gave the best AUC (0.846). These two feature selection
methods selected 11 and 29 features respectively, and the features that
were chosen in common between them were: \texttt{operation}, \texttt{bp},
\texttt{iss}, \texttt{lowerlimbnopelvis}, \texttt{age}, \texttt{lowerlimbany}
and \texttt{icu}. This adds evidence to the work of Dinh et al. who found
that \texttt{iss}, \texttt{age} and \texttt{operation} were important
distinguishers of LOS $\leq$ or $>$ 2 days \cite{Dinh2013a}. We also
confirm the findings of Gabbe et al. who found that \texttt{age},
\texttt{iss} and \texttt{bp} were useful in predicting the length of
hospital stay of blunt trauma patients \cite{Gabbe2005}. Our feature
selection results also
support the findings of McGonigal et al. \cite{McGonigal1993} who used
\texttt{age} and
\texttt{iss} to predict the probability of survival for trauma patients, a
related problem in the same medical domain.
Our observations suggest the use of feature selection not only as a
pre-processing step to improve the accuracy or AUC of a classifier, but also
as a tool used separately that could help extract insights from medical data.

Although we were able to achieve higher accuracy and AUC than the baseline,
our best approaches also have a number of limitations which need to be pointed
out. The first point to note is that using a wrapper method of feature
selection is slower than using a filter approach (or manually picking the
features before training), as the learning algorithm used as the wrapper
needs to be invoked at each of the folds of cross-validation. Additionally,
the $k$-NN algorithm becomes slower as the data set increases in the number
of examples. We did not face such a problem during our work, but this will
need to be addressed for larger data sets. It is also common to use more
than one nearest neighbour, so care needs to be taken to find a suitable number
of neighbours for a given problem. Finally, we achieved an improvement
in AUC of 0.034 but with 29 features, which is still a fairly large
number. The importance of using a fewer number of features is twofold: not
only does this speed up the training time, it also means that we are able to
make predictions without requiring a lot of information about a patient,
which is a crucial consideration if this is to be implemented for use by
physicians.

\subsection{Comparison of Classifiers}
Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
accuracy and AUC of all classifiers under each feature selection method.
The methods that do not select features based upon a specified threshold
(manual and baseline as well as C4.5 wrapper and no feature selection)
are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
based upon thresholds (correlation and information gain) are graphed
in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
threshold is changed.

In terms of classification accuracy, we did not find a lot of variation
between all classifiers, with Zero-R producing 66.91\% accuracy and 77.81\%
being the best we achieved -- an improvement of just over 10\%. The AUC of
the 11 classifiers we tested ranged from 0.498 with Zero-R to 0.846 with
logistic regression. Logistic regression, the \textit{de facto} method
used in LOS classification, performed consistently better than the others
across all feature
selection methods, but only when the feature set was not heavily reduced by
the feature selector: in these cases, such as the results from Figure
\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
selected by C4.5 decision trees and nearest
neighbour methods have higher accuracy than logistic regression. This is also
apparent from the graphs in Figure \ref{fig:tr-threshold}: for both
feature selection methods, as we increase the threshold, logistic regression
drops in performance after the initial few thresholds and is outperformed by
C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
neighbours and K*), as well as our Ranked Distance algorithm.
This is true for both accuracy and AUC.

We found that in most cases, the SVM did not achieve a higher accuracy or
AUC than logistic regression with the same features even though SVMs are
considered the state-of-the-art in classification algorithms. Additionally,
as the feature set was reduced, both the discriminating ability of SVMs
(indicated by the AUC) and accuracy decreased noticeably. SVMs have not been
used in a LOS classification problem like
ours, so there will need to be further work carried out in order to assess the
suitability of using SVMs to classify patient LOS.

@@

Nearest neighbour approaches appear to have performed well on this data set in
both accuracy and AUC, especially in comparison to more sophisticated
classifiers and when the feature set was significantly reduced. From the graphs
in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours
approaches perform similarly when the number of features is decreased, but our
Ranked Distance algorithm showed a statistically significant improvement over
other NN algorithms. This
improvement is not apparent at smaller feature sets, where using the standard
$k$-NN algorithm with 1 or 20 neighbours performs slightly better.
The performance of NN we observed in our results is
an interesting finding because nearest neighbours have not been commonly
applied in predicting the LOS, but there is intuitive appeal in the
case-by-case nature of how these algorithms classify new examples that should
be explored further. Unlike MLPs and SVMs, prediction decision of nearest
neighbours are more transparent and understandable. However, the major drawback
of NN algorithms is their storage requirement and the computational time
required to search for the nearest neighbours to make the classification.
